---
title: "Titanic: Machine Learning from Disaster"
layout: post
categories: [R]
tags: [R, Kaggle]
---

##Introduction
This blog is wrote by R Markdown.

This article following the idea of [One Approach to Deriving a Model](https://github.com/wehrley/wehrley.github.io/blob/master/SOUPTONUTS.md). The introduction of the data set is in [Titanic survival prediction challenge](https://www.kaggle.com/c/titanic-gettingStarted).

##Getting Start
Let's load this two data (train and test) into R. I use empty space to replace the missing data.
```{r, echo=T}
setwd("/media/htt/Ha/MOOC/Kaggle/Titanic")
train.column.types = c('integer', 'factor', 'factor', 'character', 'factor', 'numeric', 'integer', 'integer', 'character', 'numeric', 'character', 'factor')
train = read.csv('train.csv', colClasses = train.column.types, na.strings = c(NA, ''))
test = read.csv('test.csv', colClasses = train.column.types[-2], na.strings = c(NA, ''))
```
```{r, eval=FALSE}
train = read.csv('train.csv', colClasses = train.column.types, na.strings = c(NA, ''))
test = read.csv('test.csv', colClasses = train.column.types[-2], na.strings = c(NA, ''))
```
A view of the data set 
```{r, echo=FALSE}
head(train,n=3)
```


See the distribution of missing data in training set.
```{r, echo=T, message=F}
require(Amelia)
missmap(train,y.labels=seq(0,999,by=100),y.at=c(1,seq(0,999,by=100)), main="Missing data in training set")
```

We can see most of the Cabin data and about 20% of Age data is missing. There is no much thing we can do to Cabin data. While for Age data, we can replace the missing age with some form of imputation. 





 


Let's check the relationship between different variables:

```{r, echo=T, message=F}
barplot(table(train$Survived),
        names.arg = c("Perished", "Survived"),
        main="Survived (passenger fate)", col="black")
mosaicplot(train$Pclass ~ train$Survived, 
           main="Passenger Fate by Traveling Class", shade=FALSE, 
           color=TRUE, xlab="Pclass", ylab="Survived")
boxplot(train$Age ~ train$Pclass, main="Age and Class")

```









