---
title: "Titanic: Machine Learning from Disaster"
layout: post
categories: [R]
tags: [R, Kaggle]
---

##Introduction
This blog is wrote by R Markdown.

This article following the idea of [One Approach to Deriving a Model](https://github.com/wehrley/wehrley.github.io/blob/master/SOUPTONUTS.md). The introduction of the data set is in [Titanic survival prediction challenge](https://www.kaggle.com/c/titanic-gettingStarted).

##Getting Start
Let's load this two data (train and test) into R. I use empty space to replace the missing data.
```{r, echo=FALSE}
setwd("/media/htt/Ha/MOOC/Kaggle/Titanic")
train.column.types = c('integer', 'factor', 'factor', 'character', 'factor', 'numeric', 'integer', 'integer', 'character', 'numeric', 'character', 'factor')
train = read.csv('train.csv', colClasses = train.column.types, na.strings = c(NA, ''))
test = read.csv('test.csv', colClasses = train.column.types[-2], na.strings = c(NA, ''))
```
```{r, eval=FALSE}
train = read.csv('train.csv', colClasses = train.column.types, na.strings = c(NA, ''))
test = read.csv('test.csv', colClasses = train.column.types[-2], na.strings = c(NA, ''))
```
A view of the data set 
```{r, echo=FALSE}
head(train,n=3)
```


See the distribution of missing data in training set.
```{r, echo=FALSE}
require(Amelia)
missmap(train,y.labels=seq(0,999,by=100),y.at=c(1,seq(0,999,by=100)), main="Missing data in training set")
```






